Downstream analysis: A pipeline to analyse time-course gene expression data

### 2019-02-22 scripts/04_downstream_analysis.Rmd
ENH Better table for pathway + text

```{r echo=FALSE, results="hide"}
# Loading dependencies
library(devtools)
library(limma)
library(splines)
library(stats)
library(moanin)
library(KEGGprofile)

# library(moanin)
```

@@

@@

### Finding enriched pathways

```{r}
clusters = as.vector(unlist(unique(labels)))
for(cluster in clusters){
    genes = row.names(labels[labels == cluster])
    pathways = KEGGprofile::find_enriched_pathway(
	genes, species="mmu", download_latest=TRUE)
}
```

>>>>

```{r echo=FALSE, results="hide"}
# Loading dependencies
library(devtools)
library(limma)
library(splines)
library(stats)
library(moanin)
library(KEGGprofile)
```

@@

## Downstream analysis of clusters.

Once good clusters are obtained, the next step is to leverage the clustering
to ease interpretation. Classic enrichment analysis step can then be performed
in the gene set defined in each cluster: KEGG pathway enrichment analysis, GO
term enrichment analysis, motif enrichment analysis, etc.

@@

First, let us clean up the gene obtained and only select the genes we are
going to use in the enrichment analysis. One can either use the whole set of
genes, only the set of differentially expressed genes in each cluster, or a
subset of genes that fit well to a cluster (based on some criterion).

```{r}
gene_names = row.names(labels)
gene_names = gene_names[!is.na(labels)]
labels = labels[!is.na(labels)]
```

### Finding enriched pathways using `biomaRt` and `KEGGprofile`

Let us first tackle the case of pathway enrichment analysis. We will leverage
the packages `biomaRt` and `KEGGprofile` for this step. `KEGGprofile` is a
package that easily allows to perform pathway enrichment analysis on a set of
genes labeled with the ensembl annotation. We thus need to convert the gene
names into the appropriate format. This is where `biomaRt` comes in handy: it
enables easy conversion from one gene annotation to another. Here, we will use
it to convert the gene names from the Refseq annotation to the ensembl one.

```{r}
library(biomaRt)
ensembl = useMart("ensembl")
ensembl = useDataset("mmusculus_gene_ensembl", mart=ensembl)
```

```{r enrichment_analysis, results="hide"}
clusters = sort(as.vector(unlist(unique(labels))))
all_pathways = data.frame()
for(cluster in clusters){
    genes = gene_names[labels == cluster]

    # convert gene names
    genes = getBM(attributes=c("ensembl_gene_id", "entrezgene"),
		  filters="refseq_mrna", values=genes,
		  mart=ensembl)["entrezgene"]
    genes = as.vector(unlist(genes))
    pathways = KEGGprofile::find_enriched_pathway(
	genes, species="mmu",
	download_latest=TRUE)

    if((dim(all_pathways)[1] == 0) & (dim(pathways$stastic)[1] != 0)){
	print("Initialize")
	pathways$stastic["Cluster"] = cluster
	all_pathways = pathways$stastic
    }else if(dim(pathways$stastic)[1] != 0) {
	print("Merge")
	pathways$stastic["Cluster"] = cluster
	all_pathways = merge(all_pathways, pathways$stastic, all=TRUE)
    }
}
```

```{r echo=FALSE}
library(kableExtra)
library(viridis)

all_pathways = all_pathways[
    order(all_pathways["Cluster"],
	  all_pathways["pvalueAdj"]),]

# Select the columns to print
all_pathways = subset(
    all_pathways,
    select=c(
	"Pathway_Name", "Percentage",
	"pvalueAdj", "Cluster"))

colnames(all_pathways) = c("Pathway", "Percentage", "Adj. p-value", "Cluster")
all_pathways["Percentage"] = all_pathways["Percentage"] * 100
kable(all_pathways, "html", escape=F, digits=4) %>%
    kable_styling("striped", full_width=TRUE) 
```

### Finding enriched GO terms

```{r go_term_enrichment}

```
